---
title: "Emelie’s Violins"
author: "Benjamin Moran"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(knitr)     # This package is necessary for this type of document (.Rmd).
library(tidyverse) # This will load all of the packages we need (ggplot2, dplyr).
```

```{r options, include=FALSE}
opts_chunk$set(echo = TRUE) # prints the code chunks by default. We turn this
                            # behaviour off by calling 'echo = FALSE' in the 
                            # header of the code chunk.
```

## Setup

Here you need to import the data for each network. Store the files separately in the data folder and copy the code below once for each file. (You could do this with a loop but this is simpler for now).

### Discovery Time Data

```{r discdata}
disc.dat1 <- read.delim("data/DiscoveryTime_Random.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Seconds") # label the column

disc.dat1$Network = factor("Random") # Add a column called 'Network' with the network name

disc.dat1$Seconds = disc.dat1$Seconds/1000 # convert to seconds

disc.dat2 <- read.delim("data/DiscoveryTime_B4.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Seconds") # label the column

disc.dat2$Network = factor("B4") # Add a column called 'Network' with the network name

disc.dat2$Seconds = disc.dat2$Seconds/1000 # convert to seconds

disc.dat3 <- read.delim("data/Discovery_time_Simple3.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Seconds") # label the column

disc.dat3$Network = factor("Dim3") # Add a column called 'Network' with the network name

disc.dat3$Seconds = disc.dat3$Seconds/1000 # convert to seconds

```

Next, combine all the data files together into a single data frame.

```{r disccombinedata}
discdata = rbind(disc.dat1, # Binds the data together by row (hence 'r'bind).
             disc.dat2, disc.dat3) # add each dataset here.
```

Lastly, we need to calculate the median of each Network to plot later. Note: the `%>%` is a pipe (from the `magrittr` or `dplyr` packages in `R`). The `group_by` and `summarise` functions come from the `dplyr` package.

```{r discgetmedian}
discmed.data = discdata %>%           # Creates new object as a copy
  group_by(Network) %>%               # Groups by the network
  summarise(Median = median(Seconds)) # Returns the median value for each.

rm(disc.dat1,disc.dat2,disc.dat3)
```

### Messages Data

Here we will do the same for the messages data.

```{r messdata}
mess.dat1 <- read.delim("data/MessagesSent_B4.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Messages") # label the column

mess.dat1$Network = factor("B4") # Add a column called 'Network' with the network name
mess.dat1$Messages = mess.dat1$Messages/12 # divides B4 message numbers by 12.

mess.dat2 <- read.delim("data/Messages_sent_Simple3.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Messages") # label the column

mess.dat2$Network = factor("Dim3") # Add a column called 'Network' with the network name
mess.dat2$Messages = mess.dat2$Messages/3 # divides B4 message numbers by 12.
```

And we'll combine them in the same way.

```{r messcombinedata}
# Comment out the below line and uncomment the last lines to add extra data).
#messdata = mess.dat1

messdata = rbind(mess.dat1, # Binds the data together by row (hence 'r'bind).
                 mess.dat2) # add each dataset here.
```

And get the median.

```{r messgetmedian}
messmed.data = messdata %>%            # Creates new object as a copy
  group_by(Network) %>%                # Groups by the network
  summarise(Median = median(Messages)) # Returns the median value for each.

rm(mess.dat1,mess.dat2)
```

### Query Period Data

And for the Query period data

```{r quedata}
que.dat1 <- read.delim("data/Query_Time/Discovery_time_Period_2000.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Seconds") # label the column

que.dat1$Network = factor("2000") # Add a column called 'Network' with the network name
que.dat1$Seconds = que.dat1$Seconds/1000 # Convert to milliseconds

que.dat2 <- read.delim("data/Query_Time/Discovery_time_Period_1000.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Seconds") # label the column

que.dat2$Network = factor("1000") # Add a column called 'Network' with the network name
que.dat2$Seconds = que.dat2$Seconds/1000 # Convert to milliseconds

que.dat3 <- read.delim("data/Query_Time/Discovery_time_Period_500.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Seconds") # label the column

que.dat3$Network = factor("500") # Add a column called 'Network' with the network name
que.dat3$Seconds = que.dat3$Seconds/1000 # Convert to milliseconds

que.dat4 <- read.delim("data/Query_Time/Discovery_time_Period_200.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Seconds") # label the column

que.dat4$Network = factor("200") # Add a column called 'Network' with the network name
que.dat4$Seconds = que.dat4$Seconds/1000 # Convert to milliseconds
```

Next, combine all the data files together into a single data frame, taking the mean of every file first.

```{r quecombinedata}
# quedata = rbind(mean(que.dat1$Seconds), # Binds the data together by row (hence 'r'bind).
#             mean(que.dat2$Seconds), mean(que.dat3$Seconds), mean(que.dat4$Seconds) ) # add each dataset here.

quedata = rbind(que.dat1,que.dat2,que.dat3,que.dat4)

quedata <- quedata %>%
  group_by(Network) %>%
  summarise(meanval = mean(Seconds),
            minval  = min(Seconds),
            maxval  = max(Seconds),
            medval  = median(Seconds),
            sdval   = sd(Seconds))

rm(que.dat1,que.dat2,que.dat3,que.dat4)
```

### Ping data, lab 1

And for the ping data

```{r pingdata}
ping.dat1 <- read.delim("data/Lab1/ping_data.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Seconds") # label the column

ping.dat1$Network = factor("h1 ping h2") # Add a column called 'Network' with the network name
ping.dat1$Seconds = ping.dat1$Seconds/1000 # Convert to milliseconds

ping.dat2 <- read.delim("data/Lab1/ping_data2.txt", # fileloc
                        header = FALSE, # no header
                        col.names = "Seconds") # label the column

ping.dat2$Network = factor("h2 ping h1") # Add a column called 'Network' with the network name
ping.dat2$Seconds = ping.dat2$Seconds/1000 # Convert to milliseconds

```

Next, combine all the data files together into a single data frame.

```{r pingcombinedata}
pingdata = rbind(ping.dat1$Seconds, # Binds the data together by row (hence 'r'bind).
            ping.dat2$Seconds) # add each dataset here.
rm(ping.dat1,ping.dat2)
```


## Plotting

We'll use `ggplo2` to plot the data becuase it has an easy-to-use `geom_violin` function for violin plots. Firtsly, you need to create a ggplot object with data. Then you call the various plotting functions. Remember to plot the median dots last, because you'll need to repoint `data` at the `med.data` dataframe we created earlier.

Lastly, we'll change some of the formatting. All of these options are available in the `ggplot2` package.

```{r discplotobject}
# Create the ggplot object called 'p1'

p1 <-  ggplot(data = discdata,         # our combined data
             aes(x = Network,     # networks on the x axis
                 y = Seconds,     # Seconds on the y axis
                 group = Network) # plots each network individually
             ) +
  
      # Call each plot
      geom_violin(position = "dodge",   # Ignore
                  fill = "grey") +      # colour of violin
      geom_boxplot(width= 0.1,          # Width of the boxplot
                   outlier.colour = NA, # Outliers not plotted
                   position = "dodge",  # Ignore
                   fill = "black") +    # colour of box
      
      # plot the Median points
      geom_point(data = discmed.data,  # the median data
                 aes(x = Network,  # Networks on the x axis (same as above).
                     y = Median),  # the Median value on the y axis.
                 colour = "white", # colour of the dot
                 size = 4) +       # size of the dot
      
      # formatting options
      
      theme_linedraw() + # a basic ggplot theme
      
      labs(y = "Time (seconds)",              # y axis label
           x = "Network",                     # x axis label
           title = "Discovery Time Data") +  # Main title
      
      theme(axis.text = element_text(size=12),      # x / y axis label sizes
            axis.title = element_text(size = 18),   # x / y axis title sizes
            plot.title = element_text(size = 24,    # Main title size
                                      hjust = 0.5), # Centre main title
            panel.grid = element_blank())           # Removes all gridlines
```

And here is the plot.

### Discovery Plot

```{r discplotplot, echo = FALSE}
p1
```

\newpage 

### Messages Plots

Using the same code as the discovery time plots, we now plot the messages data.

```{r messplotobject, echo=FALSE}
# Create the ggplot object called 'p2'

p2 <-  ggplot(data = messdata,         # our combined data
             aes(x = Network,     # networks on the x axis
                 y = Messages,     # Seconds on the y axis
                 group = Network) # plots each network individually
             ) +
  
      # Call each plot
      geom_violin(position = "dodge",   # Ignore
                  fill = "grey",
                  draw_quantiles = c(0.25, 0.5, 0.75)) +  # colour of violin
      # geom_boxplot(width= 0.1,          # Width of the boxplot
      #              outlier.colour = NA, # Outliers not plotted
      #              position = "dodge",  # Ignore
      #              fill = "black") +    # colour of box
      
      # plot the Median points
      # geom_point(data = messmed.data,  # the median data
      #            aes(x = Network,  # Networks on the x axis (same as above).
      #                y = Median),  # the Median value on the y axis.
      #            colour = "white", # colour of the dot
      #            size = 4) +       # size of the dot
      
      # formatting options
      
      theme_linedraw() + # a basic ggplot theme
      
      labs(y = "Messages (#/Nodes)",              # y axis label
           x = "Network",                     # x axis label
           title = "Messages Data") +  # Main title
      
      theme(axis.text = element_text(size=12),      # x / y axis label sizes
            axis.title = element_text(size = 18),   # x / y axis title sizes
            plot.title = element_text(size = 24,    # Main title size
                                      hjust = 0.5), # Centre main title
            panel.grid = element_blank())           # Removes all gridlines
```

```{r messplotplot, echo = FALSE}
p2
```

Make a simple line plot for the query time plot

```{r queplot, echo = FALSE}
x <- c(2000, 1000, 500, 200)
plot(x, quedata, type = "o", xlab = "Query time (ms)", ylab = "Mean discovery time (s)", xlim=rev(range(x)))

 ggplot(quedata, aes(x=Network,
                     y=meanval, 
                     group = 1)) +
  geom_ribbon(aes(ymin=meanval-(1.96*sdval),
                  ymax=meanval+(1.96*sdval)),
              alpha=0.2) +
  geom_line() +
  geom_line(aes(y = medval),
            lty = 2, 
            col = "red") + 
  geom_point(aes(y = minval)) + 
  geom_point(aes(y = maxval)) + 
  ylab("")
```

And a double line plot for ping

x_ping  <- seq(1, length(ping.dat1$Seconds), 1)

plot(x_ping, log(ping.dat1$Seconds), xlab="Sequence number of echo reply", ylab="Log(RTT)", type="l",col="red")
lines(x_ping , log(ping.dat2$Seconds), col="green")
